\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-19 paper.qmd
Regenerated Figure 4 by comments

benchmark_value <- 19.4

# Filter and label small and long edges
distance_df_small_edges <- distance |>
  dplyr::filter(distance < benchmark_value) |>
  dplyr::mutate(type = "small_edges")

distance_df_long_edges <- distance |>
  dplyr::filter(distance >= benchmark_value) |>
  dplyr::mutate(type = "long_edges")

@@

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(2, 15)) |>
  dplyr::mutate(select_point = if_else(hb_id == 2, "data1", "data2"))

a1 <- ggplot(tr_df |> dplyr::filter((y == min(tr_df$y)) | (y == max(tr_df$y))), aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, color = type),
    data = tr_from_to_df_coord_with_group
  ) +
  geom_point() +
  coord_equal() +
  scale_colour_manual(values = c("#000000", "#f6e8c3"))  + 
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "#e7e1ef", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(2, 15)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == "2") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "15") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("2", "15"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::filter(hb_id %in% c("2", "15")) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://www.youtube.com/watch?v=vBvM30Plt24>.

@@

#### Benchmark value to remove low-density hexagons

@@

#### Benchmark value to remove long edges

>>>>

## Map hb_id

df_bin_n <- df_bin |>
  mutate(ID = 1:NROW(df_bin))

tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_n, by = c("from" = "ID")) |>
  rename("from_hb_id" = "hb_id")

tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_n, by = c("to" = "ID")) |>
  rename("to_hb_id" = "hb_id")

hb_id_selected <- tr_from_to_df_n |>
  dplyr::filter((from == 2) & (to == 3)) |>
  head(1)

# benchmark_value <- 19.4

# Filter and label small and long edges
distance_df_small_edges <- distance |>
  dplyr::filter((from != 2) | (to != 3)) |>
  dplyr::mutate(type = "small_edges")

distance_df_long_edges <- distance |>
  dplyr::filter((from == 2) & (to == 3)) |>
  dplyr::mutate(type = "long_edges")

@@

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))

UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))

# bin_coord_1 <- tr_df |>
#   dplyr::filter((x == hb_id_selected$x_from)) |>
#   dplyr::filter((y == hb_id_selected$y_from))

bin_coord_1 <- tr_df |>
  head(1)

# bin_coord_2 <- tr_df |>
#   dplyr::filter((x == hb_id_selected$x_to)) |>
#   dplyr::filter((y == hb_id_selected$y_to))

bin_coord_2 <- tr_df |>
  dplyr::filter(row_number() == 2)

bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)

a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to),
    data = tr_from_to_df_coord_with_group, colour = "#f6e8c3"
  ) +
  geom_point() +
  coord_equal() +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "#deebf7", aes(group = polygon_id), fill = "#ffffff") +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
    data = tr_from_to_df_coord_with_group
  ) +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
    geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_cartesian(xlim =c(-5, 8), ylim = c(-10, 10)) +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://youtu.be/yN-RNPV_cHQ>.

@@

#### Remove low-density hexagons

@@

#### Remove long edges

### 2024-01-19 paper.tex
Regenerated Figure 4 by comments

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 2nd and
15th hexagons, (b) visualization of the edge connected the 2nd and 15th
hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at
\url{https://www.youtube.com/watch?v=vBvM30Plt24}.}

@@

\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
\subsubsection{Benchmark value to remove low-density
hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}

@@

\hypertarget{benchmark-value-to-remove-long-edges}{%
\subsubsection{Benchmark value to remove long
edges}\label{benchmark-value-to-remove-long-edges}}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 2nd and
15th hexagons, (b) visualization of the edge connected the 2nd and 15th
hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at \url{https://youtu.be/yN-RNPV_cHQ}.}

@@

\hypertarget{remove-low-density-hexagons}{%
\subsubsection{Remove low-density
hexagons}\label{remove-low-density-hexagons}}

@@

\hypertarget{remove-long-edges}{%
\subsubsection{Remove long edges}\label{remove-long-edges}}
